\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2012-10-31 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
adds a (pseudo) Dirac delta source time function set by default when FORCESOLUTION is used; adds calculation of source arrays for interpolated locations of force point sources; corrects source time function factors used in GPU code; updates the user manual in Par_file and FORCESOLUTION related pages.

\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a \texttt{FORCESOLUTION} force source
instead of a \texttt{CMTSOLUTION} moment-tensor source. This can be useful e.g. for oil industry
foothills simulations in which the source is a vertical force, normal force,
tilted force, or an impact etc. By default a quasi-Heaviside time function is used to represent the force point source but a Ricker
source time function can be computed instead by changing the value of \texttt{USE\_RICKER\_IPATI} in the main constants file
\texttt{constants.h} located in the \texttt{src/shared/} subdirectory.
Note that in the \texttt{FORCESOLUTION} file, you will need to edit the East, North and vertical components of an
arbitrary (non-unitary) direction vector of the force vector; thus refer to Appendix~\ref{cha:Coordinates}
for the orientation of the reference frame. This vector is made unitary internally in the code and thus only its direction matters here;
its norm is ignored and the norm of the force used is the factor force source times the source time function.

@@

@@

\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item Set the \texttt{half duration} parameter of the source time function.
\item Set the latitude, longitude, depth of the source in geographical
  coordinates.
\item Set the magnitude of the force source.
\item Set the components of a (non-unitary) direction vector for the
  force source in the East/North/Vertical basis (see Appendix A for
  the orientation of the reference frame).
\end{itemize}

>>>>

\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a (tilted) \texttt{FORCESOLUTION} force 
point source located exactly at a grid point instead of a \texttt{CMTSOLUTION} moment-tensor source. 
When the force source does not fall exactly at a \texttt{GLL} point, the solver interpolates the force between grid points 
using Lagrange interpolants. This can be useful e.g. for oil industry foothills simulations 
in which the source is a vertical force, normal force, tilted force, or an impact etc. 
Note that in the \texttt{FORCESOLUTION} file, you will need to edit
the East, North and vertical components of an arbitrary (non-unitary)
direction vector of the force vector; thus refer to
Appendix~\ref{cha:Coordinates} for the orientation of the reference
frame. This vector is made unitary internally in the solver and thus
only its direction matters here; its norm is ignored and the norm of
the force used is the factor force source times the source time
function.
\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to 
use a Ricker source time function instead of the source time functions set by default 
to represent a (tilted) \texttt{FORCESOLUTION} force point source or a \texttt{CMTSOLUTION} moment-tensor source. 
Normally, if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo) Heaviside step 
function with a very short half duration is used for elastic cases to represent the permanent slip on the fault while 
in the acoustic case a Gaussian source time function with a similarly short half duration is used to physically describe 
actions within the fluid. Similarly, if a \texttt{FORCESOLUTION} force source is used, we need a source time function 
with finite support, so a (pseudo) Dirac delta source time function (which is actually a Gaussian of very
short half duration, i.e. hdur = 5 * DT) is used by default.
Any other source-time function may then be obtained by convolution.

@@

\item [{\texttt{GPU\_MODE}}] Turn this flag on to use GPUs.

@@

\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item Set the \texttt{half duration} parameter of the Ricker source
  time function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned
  on in the main parameter file {\texttt{Par\_file}}. In the case the
  solver uses a (pseudo) Dirac delta source time function to represent
  a force point source, a very short \texttt{half duration} parameter
  (hdur = 5 * DT) is automatically set by default.
\item Set the latitude, longitude, depth of the source in geographical
  coordinates.
\item Set the magnitude of the force source.
\item Set the components of a (non-unitary) direction vector for the
  force source in the East/North/Vertical basis (see Appendix A for
  the orientation of the reference frame).
\end{itemize}
